Metropolitan Areas (Deaths & Case)

### 2020-11-10 KPMG Team/Dashboard/msa_areas.Rmd
msa added

```{r}
# ------------------------------------------------------- #
# -----------  DATA MANIPULATION FOR CBSA --------------- # 
# ------------------------------------------------------- #

# --- *** --- core-based statistical areas  --- *** --- 
# comprised of counties 
# is a U.S. geographic area defined by the Office of Management and Budget (OMB) that consists of one or more counties (or equivalents) anchored by an urban center of at least 10,000 people plus adjacent counties that are socioeconomically tied to the urban center by commuting.
# https://en.wikipedia.org/wiki/List_of_metropolitan_statistical_areas

# select only cbsa-related information 
data_cbsa <- data %>% 
  mutate(fips = paste(`FIPS State Code`, `FIPS County Code`, sep = "")) %>%
  mutate(fips_state = `FIPS State Code`) %>%
  mutate(state = `State Name`) %>%
  mutate(fips_county = `FIPS County Code`) %>%
  rename(cbsa = `CBSA Code`) %>%
  rename(cbsa_title = `CBSA Title`) %>% 
  select(cbsa, cbsa_title, fips, state, fips_state, fips_county)

# join cbsa data and census data by GEOID/FIPS, get population (and other demographic) info.
census_cbsa_merge <- census %>% 
  rename(fips = GEOID) %>% 
  left_join(data_cbsa, by = "fips") %>% 
  group_by(cbsa) %>%
  mutate(cbsa_population = sum(population))

# -------- FILTER OR NOT? ----------- # 
# only top 12 csa and census data (run this)

# list of top cbsas by population, if we want
# top_cbsa_popn <- census_cbsa_merge %>%
#   select(cbsa, cbsa_title, cbsa_population) %>%
#   arrange(desc(cbsa_population)) %>%
#   distinct() %>%
#   na.omit() %>%
#   data.frame() %>%
#   slice(1:12)

# census_cbsa <- census_cbsa_merge %>%
#   semi_join(top_cbsa_popn, by="cbsa")

# --------------------------------------# 
# without filtering (run this)
census_cbsa <- census_cbsa_merge
# ------------------------------------ #

# only need county to merge both databsets, will keep just metro areas (csa)
cbsa_covid_merge <- census_cbsa %>% 
  left_join(nyt, by = "fips") %>% 
  select(cbsa, cbsa_title, cbsa_population, date, cases, deaths) %>% 
  filter(date >= lubridate::as_date("2020-01-24") & 
           date <= lubridate::as_date("2020-10-18")) # so that the 7 day mean wouldn't get messed up

# get 7-day mean of cases with zoo::rolling mean and dplyr::lag function 
cbsa_covid_cases <- data.frame(cbsa_covid_merge) %>% 
  group_by(cbsa, date) %>% 
  dplyr::summarise(cases_sum = sum(cases, na.rm = TRUE)) %>% 
  mutate(new_cases = cases_sum - dplyr::lag(x = cases_sum, n = 1, order_by = date)) %>%
  mutate(new_cases_mean = rollmean(new_cases, k = 7, fill = NA))

# same as above, but for deaths
cbsa_covid_deaths <- data.frame(cbsa_covid_merge) %>% 
  group_by(cbsa, date) %>% 
  dplyr::summarise(deaths_sum = sum(deaths, na.rm = TRUE)) %>% 
  mutate(new_deaths = deaths_sum - dplyr::lag(x = deaths_sum, n = 1, order_by = date)) %>%
  mutate(new_deaths_mean = rollmean(new_deaths, k = 7, fill = NA))

# join deaths and cases numbers together (not sure if I can do it in one?)
# contains
cbsa_covid <- cbsa_covid_merge %>% 
  left_join(cbsa_covid_cases, by = c("cbsa", "date")) %>% 
  left_join(cbsa_covid_deaths, by = c("cbsa", "date")) %>% 
  arrange(desc(cbsa_population))

# final count of cbsa cases, in order to get ranking
cbsa_covid_sum <- cbsa_covid %>% 
  group_by(cbsa, cbsa_title, cbsa_population) %>%
  mutate(cases = sum(cases)) %>% 
  mutate(deaths = sum(deaths)) %>%
  arrange(desc(cases)) %>% 
  select(cbsa, cbsa_title, cbsa_population, cases, deaths) %>% 
  distinct() %>% 
  na.omit()

@@

# time series data of all new cases per day (7-day rolling average) during time period specified
# names: names(cbsa_covid)
head(cbsa_covid, 10)

# Rank of CBSAs according to total cases
head(cbsa_covid_sum, 10)

@@

# combined total figures
state_covid <- nyt_state_cases %>% 
  left_join(nyt_state_deaths, by = c("state", "date")) %>%
  filter(date >= lubridate::as_date("2020-01-24") & 
         date <= lubridate::as_date("2020-10-18")

# preview
head(state_covid, 10)
```

@@

metroData <- function(cbsa_id) {
  cbsa_covid %>% 
    filter(cbsa == cbsa_id)
}

@@

@@

# Miami & Florida 
fl_covid <- stateData("Florida")
miami_covid <- metroData(33100) # from cbsa table

@@

# NYC & New York cases
ny_covid <- stateData("New York")
nyc_covid <- metroData(35620)

@@

# Dallas-Fort Worth-Arlington (19100) & Texas
tx_covid <- stateData("Texas")
dal_covid <- metroData(19100)

>>>>

```{r}
### ----- SET DATE RANGE FIRST ----- ## 
```

# ------------------------------------------------------- #
# -----------  DATA MANIPULATION FOR MSA --------------- # 
# ------------------------------------------------------- #

# --- *** --- metropolitan statistical areas  --- *** --- 
# comprised of counties 
# is a U.S. geographic area defined by the Office of Management and Budget (OMB) that consists of one or more counties (or equivalents) anchored by an urban center of at least 10,000 people plus adjacent counties that are socioeconomically tied to the urban center by commuting.
# https://en.wikipedia.org/wiki/List_of_metropolitan_statistical_areas

# select only msa-related information 
data_msa <- data %>% 
  mutate(fips = paste(`FIPS State Code`, `FIPS County Code`, sep = "")) %>%
  mutate(fips_state = `FIPS State Code`) %>%
  mutate(state = `State Name`) %>%
  mutate(fips_county = `FIPS County Code`) %>%
  rename(msa = `Metropolitan Division Code`) %>%
  rename(msa_title = `Metropolitan Division Title`) %>% 
  select(msa, msa_title, fips, state, fips_state, fips_county)

# join msa data and census data by GEOID/FIPS, get population (and other demographic) info.
census_msa_merge <- census %>% 
  rename(fips = GEOID) %>% 
  left_join(data_msa, by = "fips") %>% 
  group_by(msa) %>%
  mutate(msa_population = sum(population))

# -------- FILTER OR NOT? ----------- # 
# only top 12 msa and census data (run this)

# list of top msas by population, if we want
# top_msa_popn <- census_msa_merge %>%
#   select(msa, msa_title, msa_population) %>%
#   arrange(desc(msa_population)) %>%
#   distinct() %>%
#   na.omit() %>%
#   data.frame() %>%
#   slice(1:12)

# census_msa <- census_msa_merge %>%
#   semi_join(top_msa_popn, by="msa")

# --------------------------------------# 
# without filtering (run this)
census_msa <- census_msa_merge
# ------------------------------------ #

# only need county to merge both datasets, will keep just metro areas (csa)
msa_covid_merge <- census_msa %>% 
  left_join(nyt, by = "fips") %>% 
  select(msa, msa_title, msa_population, date, cases, deaths) %>% 
  filter(date >= lubridate::as_date("2020-01-24") & 
           date <= lubridate::as_date(Sys.Date() - 8)) # 8 days prior, so that the 7 day mean wouldn't get messed up

# get 7-day mean of cases with zoo::rolling mean and dplyr::lag function 
msa_covid_cases <- data.frame(msa_covid_merge) %>% 
  group_by(msa, date) %>% 
  dplyr::summarise(cases_sum = sum(cases, na.rm = TRUE)) %>% 
  mutate(new_cases = cases_sum - dplyr::lag(x = cases_sum, n = 1, order_by = date)) %>%
  mutate(new_cases_mean = rollmean(new_cases, k = 7, fill = NA))

# same as above, but for deaths
msa_covid_deaths <- data.frame(msa_covid_merge) %>% 
  group_by(msa, date) %>% 
  dplyr::summarise(deaths_sum = sum(deaths, na.rm = TRUE)) %>% 
  mutate(new_deaths = deaths_sum - dplyr::lag(x = deaths_sum, n = 1, order_by = date)) %>%
  mutate(new_deaths_mean = rollmean(new_deaths, k = 7, fill = NA))

# join deaths and cases numbers together (not sure if I can do it in one?)
# contains
msa_covid <- msa_covid_merge %>% 
  left_join(msa_covid_cases, by = c("msa", "date")) %>% 
  left_join(msa_covid_deaths, by = c("msa", "date")) %>% 
  arrange(desc(msa_population))

# final count of msa cases, in order to get ranking
msa_covid_sum <- msa_covid %>% 
  group_by(msa, msa_title, msa_population) %>%
  mutate(cases = sum(cases)) %>% 
  mutate(deaths = sum(deaths)) %>%
  arrange(desc(cases)) %>% 
  select(msa, msa_title, msa_population, cases, deaths) %>% 
  distinct() %>% # remove repeated values 
  na.omit()
  # drop_na(msa) # remove rows with NA in msa column

@@

# time series data of all new cases per day (7-day rolling average) during time period specified
# names: names(msa_covid)
head(msa_covid, 50)

# Rank of MSAs according to total cases
head(msa_covid_sum, 10)

@@

# combined total figures
state_covid <- nyt_state_cases %>% 
  left_join(nyt_state_deaths, by = c("state", "date")) %>%
  filter(date >= lubridate::as_date("2020-01-24") & 
         date <= lubridate::as_date(Sys.Date() - 8))

# preview
head(state_covid, 10)

@@

metroData <- function(msa_id) {
  msa_covid %>% 
    filter(msa == msa_id)
}

@@

# ---- To Find MSA codes ---- #
# head(msa_covid_sum, 10)

@@

# Miami & Florida 
fl_covid <- stateData("Florida")
miami_covid <- metroData(33124) # from msa table

@@

# NYC & New York cases
ny_covid <- stateData("New York")
nyc_covid <- metroData(35614)

@@

# Dallas-Fort Worth-Arlington (19100) & Texas
tx_covid <- stateData("Texas")
dal_covid <- metroData(19124)
